### 2015-09-17 proposal/proposal.tex
Added stub proposal.

\documentclass[10pt,a4paper]{article}

%\usepackage{fullpage}

\title{Proposal for improving content discovery by a combination linked data, natural language processing and data mining techniques}
\author{Ross Fenning}

\begin{document}

\maketitle
\thispagestyle{empty}

\section{Background}

Media companies produce ever larger numbers of articles, videos, podcasts, games, etc.
-- to which it is commonly referred as "content|.

\section{Problems}

\section{Proposal}

\section{Aims}

By the end of this proposed research, I will have:

\begin{itemize}
\item briefly considered extending this work into language other than English.
\end{itemize}

\end{document}