### 2015-12-23 report/report.tex

\documentclass[10pt,a4paper]{report}
\input{setup}

\title{Improving content discovery through combining linked data and data mining techniques}
\author{Ross Fenning}

\begin{document}

\maketitle
\tableofcontents

\input{chapters/01-intro}
\input{chapters/02-background}
\input{chapters/03-design}

\chapter{Implementation}
\label{chp:implementation}

\chapter{Evaluation}
\label{chp:evaluation}

\chapter{Conclusion}

\bibliographystyle{plain}
\bibliography{references}

\end{document}