### 2016-04-24 report/chapters/04-implementation.tex
Halfway through chapter 4

\chapter{Implementation}
\label{chp:implementation}

In this chapter, some of the details of the system implementation
are described. Initially, we discuss some of the implementation
strategy employed due the experimental nature of the system and the
desire to run multiple, competing approaches side-by-side in an
efficient way.

The remaining sections walk through the implementation of each part
of the data pipeline from upstream to feature sets and the chapter
concludes with implementation of the clustering and generation of
clusters for evaluation.

\section{Software Architecture}

All functionality was implemented as part of a single Python module
named \emph{distillery}
that is run via the Unix command line, with different subcommands for
each feature. Each command was designed around the
Unix Philosophy\cite{raymond2003art} of doing one job per command and
that they were composable via Unix pipes. This allows pipelines to
be composed:

\begin{centering}
  \begin{lstlisting}[
      basicstyle=\scriptsize,
      label=lst:unix-pipe,
      language=bash]
distillery extract <iris.txt | distillery enrich | distillery generate
  \end{lstlisting}
\end{centering}

\noindent that resemble the entire theoretical data pipeline described in
section~\ref{sec:design-pipeline} without the need for middleware
such as message queues. More typical enterprise architectures around
message oriented middleware or event-driven architecture might need
to be employed to scale this system to millions of documents, but
the simple approach above is sufficient for tens of thousands of
documents.

Even this simple approach would scale very well with suitably low
network latency. The largest bottleneck observed was the necessity
to do at least one HTTP request per item ingested and then enrich
via an HTTP request per object of a triple where the IRI is a subject.
For some content items, this could easily be over 100 HTTP requests at
which point parallel CPU cores or asynchronous I/O programming does
not overcome the demands on network performance.

Throughput is a particular challenge when data processing systems
are first launched. The day-to-day volume of item creations and
updates may be low enough for even a low-scale application, but the
performance demands increase significantly.

\section{Avoiding Redundancy Across Multiple Experiments}

The design of the data pipeline in chapter~\ref{chp:design} would
suggest an implementation with each stage implemented as a function
converting a graph into a new, richer graph.

That is, whilst the different strategies illustrated in the maximal
data pipline in Figure~\ref{fig:maximal-pipeline} can be run in
parallel, the extraction stage itself simply converts the identity
graph described in section~\ref{sec:identity-graph} to a single graph
(after having taken the union of all the strategies).

In the experiment that is the subject of this paper, the intent was
to compare each of the extraction strategies individually, but also
all possible (union) combinations thereof. This presents at least
two potential implementation strategies:

\begin{enumerate}
\item Build the whole data pipeline and machine learning system such
  that it is configurable which strategies are invoked and run an
  instance per configurable combination; or
\item implement the software to produce all possible outcomes along
  the pipeline.
\end{enumerate}

The former approach is stronger if the aim is to build the system
closer to how it would be implemented in an enterprise: only one
set of clusters is needed and even a production, enterprise system
would be configurable if maintainers wished to enable or disable
features as desired.

The latter strategy is arguably better for experimentation as every
IRI fed into the source end of the pipeline is processed at the same
time for every extraction and enrichment approach. This gives better
assurance that the same input is used to evaluate each competing
system.

Other operational reasons include being able to run the system on a
single machine if necessary and that there is no repeated work:
extraction by one means can be used in isolation, but also feed into
a union with another technique without having to recalculate that
graph. That is, if we have graphs $A$ and $B$ generated once each,
then we can output $A$, $B$ and $A \cup B$ for comparison. A parallel copy
of each system would be calculating $A$ and $B$ twice each.

The limitation here is a production-ready implementation of the
system would have be built again from the ground-up, perhaps reusing
functions and library code from the experimental version, as the
application itself will be structured around this idea of multiple
outputs for each input. This is likely to be the case for any
experimental system and is arguably in line with Brooks' classic
assertion to ``Plan to Throw One Away''\cite{brooks1995mythical}.

\begin{centering}
  \begin{lstlisting}[
      label=lst:extract-all,
      language=python,
      caption={Python function that generates all possible RDF Graphs}]
def all_extracted_graphs(iri):

    extraction_strategies = [
        dereference, extract_entities, find_links
    ]

    # This creates all 3 RDF Graphs only once
    graphs = [s(iri) for s in extraction_strategies]

    # Empty set removed from powerset
    for idx, gs in enumerate(powerset(graphs) - {{},}):
        # union defined elsewhere: list(graph) -> graph
        graph = reduce(union, gs)
        # Loop counter idx tells us from which technique
        # combination each graph comes
        yield idx, graph
  \end{lstlisting}
\end{centering}

An illustrative Python function is shown in
Listings~\ref{lst:extract-all} where each of the three extraction
techniques is invoked only once and a powerset function is used to
generate all possible unions of those graphs and thus yield all
possible usages or non-usages of each strategy to generate RDF
graphs. The details of the three extraction functions are covered in
section~\ref{impl-extraction}.

\section{Obtaining IRIs}

It is a substantial undertaking to design, build and test a
production-quality enterprise system that processing all content
produced by a media organisation. An experimental system needs to
focus on quick results and therefore is optimised to work only over
a small sample of that content.

As introduced in section~\ref{intro-problems}, all that content is
also likely to be distributed across multiple databases, content
management systems and other stores. Whilst it is hypothesised in
this research that semantics and linked data provide a good way to
extract data about the total breadth of all content without any
bespoke integration against internal systems, there still remains
the enterprise integration problem of \emph{discovery} of those
content items.

Essentially, the data pipeline outlined in
section~\ref{sec:design-pipeline} notably requires that known IRIs of
content items are fed into it, but makes no statement about how those
IRIs are acquired. This is a deliberate design decision to decouple
the concepts of discovery from this extraction/enrichment workflow.
Thus item may easily be re-ingested several times if it is updated
(or simply periodically) and
a modular approach like this defends against upstream systems changing
and having to rebuild the triggers that notify when content is created
or updated.

What follows it that the enterprise integration task has been reduced
(we do not need to write bespoke code to extract basic data about
every content item) but not wholly eliminated (some custom adaptors
need to be created to notify the pipeline of creation or update events
in each respective data store).

However creation of any such notification adaptors is out of scope
for this research altogether, so a heuristic approach was employed
where ten different, themed content aggregation pages on the BBC
website\footnote{
  Including, among others, the BBC Homepage (http://www.bbc.co.uk),
  BBC Arts (http://www.bbc.co.uk/arts) and BBC Science
  (http://www.bbc.co.uk/science)
} were polled for any new content promoted by editorial teams.
This provided a way to find new items shortly after they are published
and also ensured the data sample ultimately used focused on content
that was deemed noteworthy or interesting in the last few months.

\section{Extracting RDF Graphs}
\label{sec:impl-extraction}

\section{Enriching and not Enriching}

\section{Feature Set Generation}

\section{Feature Selection}

\section{Clustering Implementation}

\section{Results Generation Strategy}