### 2016-04-24 report/chapters/05-analysis.tex

\chapter{Results and Analysis}

\section{Comparison of Clusters Produced}

\section{Strengths and Weaknesses of Different Approaches}

\subsection{Embedded Semantics Extraction}
\subsection{Entity Extraction}
\subsection{Hyperlink Relationships}
\subsection{Enrichment by Dereference}

\section{Recommendations for Production Use}