### 2016-04-24 report/chapters/06-evaluation.tex

\chapter{Evaluation}
\label{chp:evaluation}

\section{Qualitative Survey}

\subsection{Survey Design}
\subsection{Survey Results}
\subsection{Further Observations and Discussion}

\section{Design Limitations}

\section{Implementation Difficulties}