### 2016-04-24 report/chapters/07-conclusions.tex

\chapter{Conclusions and Future Work}

\section{Challenges and Opportunitues for Semantics in the Enterprise}

\section{Suitability for Semantics and Data Mining in Media Organisations}

\section{Suggestions for Initial Implementations in the Enterprise}

\section{Future Research}
\subsection{Better Enrichment}
\subsection{Patterns for Reducing Data Noise}
\subsection{Deeper Study of Machine Learning on Semantics}