### 2016-04-24 report/report.tex

\documentclass[10pt,a4paper]{report}
\input{setup}

\title{Improving content discovery through combining linked data and data mining techniques}
\author{Ross Fenning}

\begin{document}

\maketitle
\tableofcontents

\input{chapters/01-intro}
\input{chapters/02-background}
\input{chapters/03-design}
\input{chapters/04-implementation}
\input{chapters/05-analysis}
\input{chapters/06-evaluation}
\input{chapters/07-conclusions}

\bibliographystyle{plain}
\bibliography{references}

\end{document}