### 2016-04-26 report/chapters/05-analysis.tex
Half way through chapter 5

\chapter{Results and Analysis}

In chapter~\ref{chp:design}, a data pipeline is derived that -- at
the logical maximum (see figure~\ref{fig:maximal-pipeline}) could
generate feature sets using any of 112 combinations of extraction
and enrichment techniques. This is based on the product of using or
not using any of three described extraction approaches (less the case
where no extraction is performed) and enabled or disabling any of
four distinct enrichment approaches (retaining the case where no
enrichment is performed).

More practically, in chapter~\ref{chp:implementation}, the
implementation of a real system for evaluating a subset of only
fourteen combinations of approaches is described. Therefore there
are fourteen resultant sets of clusters of the same BBC web content
produced for comparative analysis.

In the following sections of this chapter, an objective and technical
analysis of the clusters is performed, leading to a critical review
of each individual extraction and enrichment approach, based on how
each respective technique appears to contribute to the results.

In chapter~\ref{chp:evaluation}, there is an evaluation of the
overall system based
on qualitative data gathered by a survey of real human users. That
evaluation is more focused on the suitability of the holistic use of
semantics and machine learning in the real world application of
suggesting related media content to web users.

Conversely, in this chapter, we take a much narrower focus on the
shape, size and quality of clusters themselves.

\section{Brief Comparison of Clusters Produced}

The most stark difference between the results produced is the large
range in terms of number of clusters produced for each approach
respectively.

\begin{table}[h]
  \centering
  \caption{Number of clusters produced by each combination of approaches}
  \label{tbl:cluster-counts}
  \begin{tabular}{lll}
    & Not enriched & Enriched \\
    Deference                                     & 5            & 249      \\
    Entity Extraction                             & 35           & 32       \\
    Hyperlink Relationships                       & 2            & 56       \\
    Dereference and Entity Extraction             & 329          & 340      \\
    Entity Extraction and Hyperlink Relationships & 199          & 233      \\
    Dereference and Hyperlink Relationships       & 53           & 117      \\
    All three extractions                         & 184          & 223     
  \end{tabular}
\end{table}

Table~\ref{tbl:cluster-counts} shows the number of clusters produced
by each combination of techniques. The range here is explained by
some variants of the system produced few, very large clusters and
some a large number of smaller clusters.

Which of the two outcomes is more suitable is arguably a matter for
real evaluation, but it is perhaps reasonable to claim that clusters
which are too large are less useful for choosing related content items
to suggest to users. In the case of the unenriched graphs obtained
via hyperlink relationships only, one of the two clusters produced
contains 2,323 items -- a number that risks ``related content''
suggestions being too erratic and random.

A deeper inspection of that largest cluster reveals every item was
deemed to be related to an FAQ page on how to share BBC pages on
social media websites -- a page that appears to be linked to from
every BBC programme information page or iPlayer catch-up video page.
This is clearly not a good indication of the content of the page
and raises two immediate issues:

\begin{enumerate}
\item the system would benefit from further heuristics that avoid
  inferring too many relationships from ancillary pages such as FAQs
  that are unrelated to the content itself; and
\item the nature of agglomerative, hierarchical clustering means
  that this system variant spent a large amount of time grouping all
  pages that link to this FAQ page at the expense of other,
  potentially useful clusters. This is a consequence of hierarchical
  clustering being a greedy algorithm.
\end{enumerate}

Further discussion of the merits and problems with hyperlink
relationships is covered in section~\ref{sec:anal-hyperlink}.

If we look at the other end of the spectrum, the system variant
that produced the most clusters was the enriched union of dereference
and entity extraction. With so many clusters, the distribution of
sizes is better shown by the histogram in figure~\ref{fig:sizes-3-1}.
Here we can see the results are dominated by clusters with 25 or
fewer items.

\begin{sidewaysfigure}
  \begin{tikzpicture}[xscale=3,yscale=2]
    \begin{axis}[
        ybar,
        ymin=0
      ]
      \addplot +[
        hist={
          bins=12,
          data min=0,
          data max=300
        }   
      ] table [y index=0] {../data/sizes-3-1.csv};
    \end{axis}
  \end{tikzpicture}
  \caption{Distribution of cluster sizes produced by the enriched union of dereference and entity extraction}
  \label{fig:sizes-3-1}
\end{sidewaysfigure}

An exploratory insight into the makeup of these clusters is shown
in table~\ref{tbl:top-features-3-1} in which the ten most common
features are shown for the one of the smallest, one of the biggest
and one of the mean-sized clusters.

\begin{sidewaystable}[p]
  \centering
  \begin{tabular}{|l|l|l|}
    $N=2$                                 & $N=9$                                                   & $N=265$                                     \\
    \hline
    ogp:image="[\ldots]"                  & rdf:type\_schema:RadioEpisode=true                      & md:item\_rdf:nil=true                       \\
    foaf:topic\_dbpedia:BBC\_iPlayer=true & schema:url="[\ldots]"                                   & foaf:topic\_dbpedia:Digg                    \\
    rdf:type\_schema:RadioEpisode=true    & foaf:topic\_dbpedia:JavaScript                          & twitter:card:"summary\_large\_image"        \\
    rdfa:usesVocabulary\_schema=true      & meta:twitter:card="summary"                             & foaf:topic\_dbpedia:LinkedIn                \\
    meta:twitter:card="summary"           & foaf:topic\_dbpedia:Listen\_(Beyonc\'e\_Knowles\_song)" & meta:apple-mobile-web-app-title="BBC Sport" \\
  \end{tabular}
  \caption{Ten most common features across the smallest, the largest and a mean-sized cluster produced by the enriched union of dereference and entity extraction}
  \label{tbl:top-features-3-1}
\end{sidewaystable}

The smallest cluster with 2 items in table~\ref{tbl:top-features-3-1}
seems to contain two pages that mention BBC iPlayer (the
\texttt{foaf:topic} predicate suggests this was found via entity
extraction) and whose primary content is an item with \texttt{rdf:type}
``RadioEpisode'' in the Schema.org\footnote{http://schema.org/}
vocabulary. This shows hints of a strong complement between semantic
web data and entity extraction confirming that textual content on the
page correlates with any semantics also embedded in the page.

Interestingly, this cluster is also strengthened by the fact that
both pages have a common thumbnail (the \texttt{ogp:image} predicate
indicates a thumbnail when used to share on Facebook). An examination
of the two content pages in question confirms that they are two
episodes of the same ``5 Live Breakfast'' programme on the
\emph{BBC Radio 5 live} radio channel.

The other two predicates for the small cluster are less clear without
some explanation as to their origin. The \texttt{meta:twitter:card}
predicate is a
documented\footnote{https://dev.twitter.com/cards/types/summary}
HTML meta tag that to indicate a page has included further meta tags
to control which metadata to include in the so-called
\emph{Summary Cards} Twitter generates as a summary of page a user
is sharing in a post to the website. This overlaps some of the
features provides by the \emph{Open Graph protocol}\footenote{http://ogp.me/}
vocabulary used by Facebook for similar means.

This overlap between Facebook and Twitter highlights both:

\begin{itemize}
  \item a failure
    for industry to adopt (or for academia to convince them to adopt)
    semantic web initiatives such as RDFa,
    Microdata, Microformats or Schema.org -- designed to provide
    machine-readable semantics in ways richer than simple meta tags; but
    also
  \item a critical opportunity for research and development in semantics
    to build on top of the real, commercial pressure created by Facebook,
    Twitter, Google et al to enrich pages with semantic data so that
    content providers can control how their content is summarised when
    shared on their platforms and also how their content is analysed for
    searching and extrapolation of real time trends in social media
    conversation.
\end{itemize}

That predicates for Facebook and Twitter are appearing in this
experiment so easily also shows evidence that the BBC is among the
media organisations that now have content producers providing
semantic enrichment alongside their content production or at least
that automated generation of in-page semantics from metadata already
present in content management systems is being prioritised alongside
other work to develop features more visible to human users. It is not
unreasonable to suppose this drive is at least partly driven by
content producers' and content promoters' desire to control how
content is displayed and discovered on third party social media
platforms.

Returning to the mean-sized cluster in table~\ref{tbl:top-features-3-1},
we see a cluster where the most common items are identified as pages
each for a single episode. There is no indication in the top five
features that they are episodes of the \emph{same} programme, however.
This could well be a strength of this system variant in that it is
able to suggest other radio programmes on the basis it has understood
they all share a common type, but it is open to suggesting episodes
of other programmes for greater diversity. In fact, this may be a
necessary feature in any ``related content'' feature as it could
be safely assumed that any BBC page for a single episode of a radio
programme already contains clear and sufficient navigation links to
find more episodes of the same programme; machine learning adds no
value to the user here.

Again, there is potential to call out a real benefit in the
combination of in-page semantics for this ``hard'' categorical
data (i.e. the page's content is of type ``Radio Episode'') with
entity extraction for a looser, schemaless vocabulary of words that
more qualitatively describe the topics covered in the content.

In this case, the topics extracted appear to be the JavaScript
programming language and the song \emph{Listen} by music artist
Beyonc\'e Knowles. A manual review of the pages involved finds them
to be radio episode pages again, but with no clear indication that
they mention either topic. Deeper analysis of the HTML source for
these pages finds two ``mistakes'' the system has made:

\begin{itemize}
\item the word JavaScript appears in the page only as part of the
  MIME type \texttt{text/javascript} in markup around JavaScript
  \emph{code} supporting the page; and
\item the word Listen appears on \emph{every} BBC page concerning
  a radio programme due to the fact they all contain a ``Listen''
  button allowing the user to listen to the live radio stream for
  that particular radio station. DBpedia Spotlight has assumed this
  word is a mention of a song titled \emph{Listen}.
\end{itemize}

\section{Deeper Analysis of Clusters}

\section{Strengths and Weaknesses of Different Approaches}

\subsection{Embedded Semantics Extraction}
\subsection{Entity Extraction}
\subsection{Hyperlink Relationships}
\label{sec:anal-hyperlink}
\subsection{Enrichment by Dereference}

\section{Recommendations for Production Use}