### 2016-04-27 paper/paper.tex
Updated abstract

\documentclass{sig-alternate-05-2015}

\begin{document}

\title{Combining Linked data and Data Mining Techniques to improve Clustering of Large Scale Media Repositories: A case study with BBC}
%\subtitle{[Extended Abstract]

\numberofauthors{3}
\author{
\alignauthor
Ross Fenning\\
       \affaddr{British Broadcasting Corporation (BBC)}\\
       \affaddr{MediaCityUK}\\
       \affaddr{Salford, United Kingdom}\\
       \email{ross.fenning@bbc.co.uk}
\alignauthor
Dhavalkumar Thakker\\
       \affaddr{University of Bradford}\\
       \affaddr{Bradford}\\
       \affaddr{United Kingdom}\\
       \email{dhavalkumar.thakker@gmail.com}
\alignauthor Tejal Shah\\
       \affaddr{UNSW}\\
       \affaddr{Australia}\\
       \email{tejals@cse.unsw.edu.au}
}

\maketitle
\begin{abstract}
Media companies produce ever larger numbers of articles, videos, podcasts,
games, commonly collectively known as ``content''. A successful
content-producing organisation not only has to develop systems to aid producing
and publishing content, but there are also demands to engineer effective
mechanisms to aid consumers in finding that content. Linked Data technologies
provide data enrichment beyond attributes and keywords explicitly available
within content data or metadata. In our work we experiment with a
plausible mapping of an RDF graph representing a content item to an attribute
set suitable for data mining. With such a mapping, we have explored the
application of machine learning, particularly unsupervised learning, across
an organisation's whole content corpus. We have created an innovative pipeline
of linked data and data mining that allows clustering of media content items on
the fly. We have evaluated such system in the context of website content
produced by the British Broadcasting Corporation (BBC) and present the optimum
combination of RDF graphs and data mining with qualitative and quantitative
results.
\end{abstract}

\keywords{Semantics; Linked Data; Semantic Web; Machine Learning; Clustering; Data Mining; RDF Graph; Media repositories; Content Management}

\section{Introduction}

\section{Background}

\section{Design}

\section{Implementation}

\section{Analysis}

\section{Evaluation}

\section{Conclusions}

%\bibliographystyle{abbrv}
%\bibliography{../report/references}

\end{document}